---
title: "Homework 4"
author: "Zachary Lazerick"
date: "7 October, 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Problem 3
#Analysis of the Pharmacy Data

a) Read in the Data

```{r}
Pharmacy = read.csv("Pharmacy.csv")
attach(Pharmacy)
```

b) Fit a Linear Model with Predictor "Ingredients Purchase Directly" and Response "Sales Volume"

```{r}
Pharmacy.fit <- lm(Sales.Volume~Ingredients.Purchased.Directly, data=Pharmacy)
summary(Pharmacy.fit)
```

c) Confirm Estimates for Beta_0 and Beta_1 from Problem 1

```{r}
plot(Ingredients.Purchased.Directly, Sales.Volume)
abline(Pharmacy.fit, lwd = 3, col = "red")
coef(Pharmacy.fit)
```
These reported coefficients match the hand-calculated coefficients in Problem 1.

d) Use Output to Report the Estimate for V(e) = sigma^2

The output reports a Residual Std. Error of 9.022 on 8 Degrees of Freedom. Squaring this result gives the estimate for V(e), so (9.022)^2 = 81.39648.

e) How are the 8 Degrees of Freedom Obtained? 

There are 8 Degrees of Freedom because the sample size, n = 10 and the formula for Degrees of Freedom for Simple Linear Regression is d.f. = n - 2, because we estimate Beta_0 and Beta_1. Substituting 10 for n yields 10 - 2 = 8 d.f.

##Problem 4
#Analysis of the Casino/Crime Data

a) Read in the Data

```{r}
Casino = read.csv("Casino.csv")
attach(Casino)
```

b) Fit a Linear Model with Predictor "Number of Casino Employees" and Response "Crime Rate"

```{r}
Casino.fit <- lm(Crime.Rate~Number.of.Casino.Employees, data=Casino)
summary(Casino.fit)
```

c) Plot the Data along the Regression Line

```{r}
plot(Number.of.Casino.Employees, Crime.Rate)
abline(Casino.fit, lwd = 3, col="blue")
```

d) Use the Output to Report the Estimate for V(e) = sigma^2

The output reports a Residual Std. Error of 0.3446 on 8 Degrees of Freedom. Squaring this result gives the estimate for V(e) = (0.3446)^2 = 0.1187492.

e) Predict the Crime Rate when there are 25,000 Casino Employees

Given that there are 25,000 Casino Employees, this means that the 'Number.of.Casino.Employees' is 25. So, substituting 25 for X in our regression equation, we get Y = Beta_0 + Beta_1 * X = (-.86977) + (0.11493)(25) = 2.09896. Thus the expected Crime Rate when there are 25,000 Casino Employees is 2.098986 (per 1,000 population).

