---
title: "HW6Q1"
author: "Zachary Lazerick"
date: "11/4/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r global.options, include=FALSE}
knitr::opts_chunk$set(
  fig.align = 'center'
)
```

Load the Lawstat Library to Access the Brown-Forscythe-Levene Test
```{r}
library(lawstat)
```

Read in the Data
```{r}
Advertising = read.csv("Advertising.csv")
attach(Advertising)
```

**Question 1: Building an SLR Model Sales ~ Radio**
(Initial Results)

Construct a Simple Linear Model with Predictor: Radio and Response: Sales
```{r}
lm.fit.radio = lm(sales~radio, data = Advertising)
```
```{r, echo=F}
summary(lm.fit.radio)
```

Construct a Scatter Plot with the Calculated Linear Model
"Eye-Test" for Abnormalities (Non-linearity, etc.)
```{r, echo=F}
par(mfrow=c(1,1))
plot(radio, sales)
abline(lm.fit.radio, col = 'red', lwd = 3)
```

Construct Plots to Check Diagnostics
```{r, echo=F}
plot(lm.fit.radio, which = c(1))
plot(lm.fit.radio, which = c(2))
plot(lm.fit.radio, which = c(5))
```

Check Distribution of Residuals for Obvious Deviations
```{r, echo=F}
boxplot(lm.fit.radio$residuals, ylab="Residuals")
hist(lm.fit.radio$residuals)
```

Run Shapiro-Wilk and Brown-Forscythe-Levene Tests
```{r}
shapiro.test(lm.fit.radio$residuals)
```
```{r, echo=F}
Advertising$Group = rep("Group1", 200)
indexRadio = (radio > median(radio))
Advertising$Group[indexRadio] = "Group2"
```
```{r}
levene.test(lm.fit.radio$residuals, Advertising$Group, location = c("median"))
```

**Conclusion**
From the Shapiro-Wilks Test, we conclude that the data is not from
a normal distribution, and from the Levene Test, we conclude that the data is
not homoscedastic. Therefore, we must transform the data
in hopes of obtaining approximately normal and homoscedastic data to draw
statistically significant conclusions from this data.

**Data Transformation Test Results**

|   **Response ~ Predictor**  	| **Shapiro-Wilks P-Value** 	| **Levene P-Value** 	|
|:-----------------------:	|:---------------------:	|:------------------------------:	|
|      Sales ~ Radio      	|       2.367e-05       	|            1.671e-10           	|
|   sqrt(Sales) ~ Radio   	|       3.443e-09       	|            4.465e-06           	|
|   log10(Sales) ~ Radio  	|        7.05e-14       	|             .003309            	|
|    (1/Sales) ~ Radio    	|        2.2e-16        	|              .3412             	|
| (1/Sales) ~ sqrt(Radio) 	|        2.2e-16        	|              .2836             	|

**Conclusion**
Upon transforming our Response Variable Sales to (1/Sales), we are able to obtain homoscedastic data. However, we are unable to transform the corresponding Predictor Radio, to obtain  approximately normal data in conjunction with (1/Sales) to draw statistically significant data because one of the levels of Radio is 0. Therefore, transformations like (log10Radio) and (1/Radio) yield undefined values, and thus a simple linear regression model, with Shapiro-Wilks and Levene test results, cannot be obtained from these transformations. 
